---
title: "Steps for running the analysis"
format: html
---


To reproduce the results from this simulation, run the following steps. All model runs were run on the Open Science Grid (OSG), and an account is necessary use it. 

See `steps_for_OM_EM_setup.qmd` for how recruitment deviation and F matrices were created for the OMs. Once all necessary inputs are created, create directories for each scenario and tar them to prepare them for uploading to OSG. The `Inputs` directory should contain: 

*   poor_recdevs_mat.RData
*   recdevs_mat.RData
*   setseed.csv
*   sas.csv  
*   effN.csv
*   F_df.csv  

```{r}
library(fs)
load(file.path(main.dir, "good_its.RData"))

scen <- "IRFSelex_poorrec_qprior"
nyears_fwd <- 25
niter <- 100
scen_dirs <- file.path(
    main.dir,
    paste(scen, nyears_fwd, "yrfwd", sep = "_"),
    good_its_100
)
sapply(scen_dirs, dir.create, recursive = T)

if(grepl("poorrec", scen)){
    file.copy(from = file.path(main.dir, "Inputs", "poor_recdevs_mat.RData"), file.path(scen_dirs, "poor_recdevs_mat.RData"))
}

file.copy(
    from = file.path(main.dir, "Inputs", "setseed.csv"),
    file.path(scen_dirs, "setseed.csv")
)
file.copy(
    from = file.path(main.dir, "Inputs", "sas.csv"),
    file.path(scen_dirs, "sas.csv")
)
file.copy(
    from = file.path(main.dir, "Inputs", "constantF_mat.RData"),
    file.path(scen_dirs, "constantF_mat.RData")
)
file.copy(
    from = file.path(main.dir, "Inputs", "effN.csv"),
    file.path(scen_dirs, "effN.csv")
)


for (i in 1:length(scen_dirs)) {
    fs::dir_copy(
        file.path(
            main.dir,
            "models",
            paste0("opaka-om-", nyears_fwd, "-selex-qprior")
        ),
        scen_dirs[i]
    ) #"-FRS-only" #, "-selex" , "-r0trend", "_FRSlencomp"
    fs::dir_copy(
        file.path(
            main.dir,
            "models",
            paste0("opaka-em-", nyears_fwd, "-selex-qprior")
        ),
        scen_dirs[i]
    )
}

# only for FRS_only_poorrec_25_yrfwd scenario (testing hyperstablity)
if(scen == "FRSonly_poorrec"){
    for(i in 1:length(scen_dirs)){
        file.copy(from = file.path(main.dir, "FRSonly_SQ_25_yrfwd", i, "em", "ss3.dat"),
        to = file.path(scen_dirs[i], "normal_rec.dat"))
    }
}

for (i in 1:length(scen_dirs)) {
    shell(paste0("powershell cd ", scen_dirs[i], ";tar -czf start.tar.gz ./*"))
    files_clean <- list.files(
        scen_dirs[i],
        pattern = ".csv|.RData|opaka-|normal_rec.dat",
        full.names = T
    )
    unlink(files_clean, recursive = T)
}

target_dirs <- paste0(paste(scen, nyears_fwd, "yrfwd", sep = "_"), "/", good_its_100)
writeLines(
    target_dirs,
    file.path(
        main.dir,
        paste(scen, nyears_fwd, "yrfwd", sep = "_"),
        "target_dirs.txt"
    )
)
shell(paste0(
    "powershell cd ",
    main.dir,
    ";tar -czf upload.",
    paste(scen, nyears_fwd, "yrfwd", sep = "_"),
    ".tar.gz ",
    paste(scen, nyears_fwd, "yrfwd", sep = "_")
))
#upload these files to OSG
```

```{r}
shell(paste0(
    "powershell cd ",
    main.dir,
    ";tar -czf scripts.tar.gz ",
    "scripts"
))
```

```{bash}
#log in to OSG in powershell
ssh meg_oshima@ap41.uw.osg-htc.org
# put in password when prompted
```


```{bash}
#in separate powershell termnial, upload files 
# sceneario folder
scp upload.HRF_poorrec_25_yrfwd.tar.gz meg_oshima@ap41.uw.osg-htc.org:/home/meg_oshima/ 
# scripts 
scp scripts.tar.gz meg_oshima@ap41.uw.osg-htc.org:/home/meg_oshima/ 
# build container 
scp R/linux-r4ss-v2.def meg_oshima@ap41.uw.osg-htc.org:/home/meg_oshima/singularity/linux_r4ss 
```


```{bash}
tar -xzf scripts.tar.gz
chmod 777 scripts/osg-prep-HRF-poorrec.sh
tar -xzf upload.HRF_poorrec_25_yrfwd.tar.gz HRF_poorrec_25_yrfwd
./scripts/osg-prep-HRF-poorrec.sh
condor_submit scripts/condor_create_OMEM_HRF_poorrec.sub
```


```{bash}
# Download results 
scp meg_oshima@ap41.uw.osg-htc.org:/home/meg_oshima/download.HRF_poorrec_25_yrfwd.tar.gz ./download.HRF_poorrec_25_yrfwd.tar.gz

```




```{r}

test_its <- c(87,66,48,59, 1, 36, 30, 89, 81)
ts_re_full %>%
filter(year <= 2048 & scenario == "IRFSelex_poorrec_25_yrfwd") %>%
filter(iteration %in% test_its) %>% 
mutate(it_cat = case_when(
    iteration %in% test_its[1:4] ~ "bad",
    iteration %in% test_its[5:9] ~ "good"
)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
#geom_ribbon(aes(x = year, ymin = lci, ymax = uci, fill = scen), alpha = .25) + 
geom_line(aes(x = year, y = SpawnBio_re, color = it_cat, group = iteration), linewidth = 1.2) + #%in% test_its[1:4]
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "SSB MRE") + 
theme_bw()
```